---
title: "Assignment# 3"
output: html_document
---

```{r setup, include=FALSE}
library(corrplot)
knitr::opts_chunk$set(echo = TRUE)
eval_data_url <- 'https://raw.githubusercontent.com/jjohn81/DATA621_Assignment_3/master/crime-evaluation-data_modified.csv'
train_data_url <- 'https://raw.githubusercontent.com/jjohn81/DATA621_Assignment_3/master/crime-training-data_modified.csv'
```

### Import Data


```{r Import}
train_data <- read.csv(train_data_url)
eval_data <- read.csv(eval_data_url)

```
  
    Training dataset contains `r nrow(train_data)` observations and `r ncol(train_data)` variables.
    Evaluation dataset contains `r nrow(eval_data)` observations and `r ncol(eval_data)` variables.
    There are no missing values in the dataset; however, there are alot of zero values( fo example, zn         
    variable) and we will treat those values as valid.

#### Summary 
    I dont know how to summarize this data. Just printing summary or boxplot doesnt really explain the data well. 
    From histograms and QQ plots, we see the data is not normally distributed and skewed.  
    
    #TODO
        Transformations?
    

```{r}
 

par(mfrow=c(3, 3))
colnames <- dimnames(train_data)[[2]]

  for(col in 2:ncol(train_data)) {
hist(train_data[,col],breaks = 25)
qqnorm(train_data[,col])
qqline(train_data[,col])
    
    
  }

```

```{r}
train <- as.data.frame((train))

par(mfrow=c(3, 3))
colnames <- dimnames(train)[[2]]

  for(col in 2:ncol(train)) {

    d <- density(na.omit(train[,col]))
   #d <- qqnorm(na.omit(train[,col]))
    plot(d, type="n", main=colnames[col])
    polygon(d, col="blue", border="gray")
  }
```

#### cor plot
```{r}
#TODO -- Explain this corr 
corrplot(cor(train_data))
corrplot(cor(train_data), order = "AOE")
```


#### Data prep

    We might need to do some sort of transformation. 
    
    
#### Model Selection
    
