---
title: "Assignment# 3"
output: html_document
---

```{r setup, include=FALSE}
library(corrplot)
library(caret)
library(ggplot2)
library(pROC)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
eval_data_url <- 'https://raw.githubusercontent.com/jjohn81/DATA621_Assignment_3/master/crime-evaluation-data_modified.csv'
train_data_url <- 'https://raw.githubusercontent.com/jjohn81/DATA621_Assignment_3/master/crime-training-data_modified.csv'
```

### Import Data


```{r Import}
train_data <- read.csv(train_data_url)

set.seed(101) # Set Seed so that same sample can be reproduced in future also
# Now Selecting 75% of data as sample from total 'n' rows of the data 
##https://stackoverflow.com/questions/17200114/how-to-split-data-into-training-testing-sets-using-sample-function
sample <- sample.int(n = nrow(train_data), size = floor(.75*nrow(train_data)), replace = F)
train <- train_data[sample, ]
test  <- train_data[-sample, ]

eval_data <- read.csv(eval_data_url)

```
  
    Training dataset contains `r nrow(train_data)` observations and `r ncol(train_data)` variables.
    Evaluation dataset contains `r nrow(eval_data)` observations and `r ncol(eval_data)` variables.
    There are no missing values in the dataset; however, there are alot of zero values( fo example, zn         
    variable) and we will treat those values as valid.

#### Summary 
    I dont know how to summarize this data. Just printing summary or boxplot doesnt really explain the data well. 
    From histograms and QQ plots, we see the data is not normally distributed and skewed.  
    
    #TODO
        No need to transformation since we are using GLM 
        No need to look for normality and diagnostic plots either
    

```{r}
 

#par(mfrow=c(3, 3))
#colnames <- dimnames(train_data)[[2]]

#  for(col in 2:ncol(train_data)) {
#hist(train_data[,col],breaks = 25)
#qqnorm(train_data[,col])
#qqline(train_data[,col])
    
    
#  }

```

```{r}
train <- as.data.frame((train_data))

#par(mfrow=c(3, 3))
#colnames <- dimnames(train)[[2]]

 # for(col in 2:ncol(train)) {

  #  d <- density(na.omit(train[,col]))
   #d <- qqnorm(na.omit(train[,col]))
   # plot(d, type="n", main=colnames[col])
  #  polygon(d, col="blue", border="gray")
  #}
```

#### cor plot
variables 'rm' and 'medv'  have vif above 5 and raises concerns. 
```{r}
#cor(train_data)
#TODO -- Explain this corr 
corrplot(cor(train_data))
full.model <- glm(train_data$target ~ ., family=binomial, train_data)

car::vif(full.model)
```


#### Data prep

    We might need to do some sort of transformation. 
    
    
#### Model Selection
#### look into pca as well here. 
We are using backward stepwise selection to choose models. We will begint with full/saturated model and eliminate any variables that are not statistically signinficant. 

#### Full Model

  Contains all the predictor variables.

```{r}
full.model <- glm(target ~ ., family=binomial, train)
summary(full.model)

```
#### Second Model
   This model contains only statistically significant variables from the full model; only varibles with p-value less than  .05 is included in this model. 

```{r}
sig.model <- glm(target ~  nox + age + dis + rad + tax + ptratio + medv ,data=train, family=binomial)
summary(sig.model )

```

#### Third Model Model
   This model excludes any varibles with high VIF.
```{r}
low.vif.model <- glm(train_data$target ~  zn+indus+chas+nox+age+dis+rad+tax+ptratio+lstat     
, family=binomial, train)
summary(low.vif.model)

```
#### PCA Model
 This model excludes any varibles with high VIF.
```{r}
pca.model <- glm(train_data$target ~  zn+indus+chas+nox+age+dis+rad+tax+ptratio+lstat     
, family=binomial, train)
summary(pca.model )

```

#### Model Selection
    Do these for all three models and select a model based on these and explain.
```{r}
pca <- prcomp(test)
summary(pca)
round(pca$rot[,1],1)
plot(pca$sdev[1:12],type="l",ylab="SD of PC", xlab="PC number")
which.min(pca$val)
```
### Prediction. 
```{r}
full.predict <- predict(full.model, newdata = eval_data, type="response")
full.predict  <- ifelse(full.predict  > 0.5,1,0)

pca.eval.data <- eval_data %>% 
    select(zn, indus, chas, nox, age, dis, rad, tax, ptratio,lstat)
pca.model.predict <- predict(pca.model, newdata = pca.eval.data, type="response")
pca.model.predict  <- ifelse(pca.model.predict  > 0.5,1,0)

low.vif.eval.data <- eval_data %>% 
    select(zn, indus, chas, nox, age, dis, rad, tax, ptratio, lstat)

low.vif.model.predict <- predict(low.vif.model, newdata = low.vif.eval.data, type="response")
low.vif.model.predict  <- ifelse(low.vif.model.predict  > 0.5,1,0)

sig.eval.data <- eval_data %>% 
    select( nox  , age  , dis  , rad  , tax  , ptratio  ,medv)


#sig.model.predict <- predict(sig.model, newdata = sig.eval.data, type="response")
#sig.model.predict  <- ifelse(sig.model.predict  > 0.5,1,0)

#pca.predict <- predict(pca.model, newdata = eval_data)
#low.vif.predict <- predict(low.vif.model, newdata = eval_data)
#sig.predict <- predict(sig.model, newdata = eval_data)


```
##confusion matrix
Confussion matrix for full model
```{r}
full.predict <- predict(full.model, newdata = test, type="response")
full.predict.target  <- ifelse(full.predict  > 0.5,1,0)
names(full.predict) <- c("target")
confusionMatrix(table(full.predict.target, test$target))

```

```{r}
sig.model.predict <- predict(sig.model, newdata = test, type="response")
sig.model.predict.target  <- ifelse(sig.model.predict  > 0.5,1,0)
names(sig.model.predict.target) <- c("target")
confusionMatrix(table(sig.model.predict.target, test$target))

```
```{r}
pca.model.predict <- predict(pca.model, newdata = test, type="response")
pca.model.predict.target  <- ifelse(pca.model.predict  > 0.5,1,0)
names(pca.model.predict.target) <- c("target")
confusionMatrix(table(pca.model.predict.target, test$target))

```

```{r}
low.vif.model.predict <- predict(low.vif.model, newdata = test, type="response")
low.vif.model.predict.target  <- ifelse(low.vif.model.predict  > 0.5,1,0)
names(low.vif.model.predict.target) <- c("target")
confusionMatrix(table(low.vif.model.predict.target, test$target))

```
##ROC and AUC
```{r}
rocCurve <- roc(test$target, full.predict, levels=c(0,1))

plot(rocCurve, legacy.axes = TRUE)
auc(rocCurve)
```
```{r}
rocCurve <- roc(test$target, sig.model.predict, levels=c(0,1))

plot(rocCurve, legacy.axes = TRUE)
auc(rocCurve)
```

```{r}
rocCurve <- roc(test$target, pca.model.predict, levels=c(0,1))

plot(rocCurve, legacy.axes = TRUE)
auc(rocCurve)
```

```{r}
rocCurve <- roc(test$target, low.vif.model.predict, levels=c(0,1))

plot(rocCurve, legacy.axes = TRUE)
auc(rocCurve)
```

##AIC 
```{r}
full.model$aic
sig.model$aic
pca.model$aic
low.vif.model$aic
```
## deviance
```{r}
anova(full.model)
anova(sig.model)
anova(pca.model)
anova(low.vif.model)
```

